---
title: '' output: html_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  include = FALSE
)
so <- function() {
  source('src.r')
  library(cowplot)
}
so()

## Theme ----
# don't know how to integrate this into a theme() call
colors    <- c('#BC0722', '#596f91')
go_colors <- c(colors, '#f16913', '#0DAC4B')
colors_ls <- list(
  disease        = colors[[1]],
  recovery       = colors[[2]],
  protein        = go_colors[[3]],
  prot_disease   = go_colors[[3]],
  prot_recovery  = go_colors[[4]],
  mot_blue = get_palette2(3, 'Blues', 4)[1],
  pro_oran = get_palette2(3, 'Oranges', 4)[1],
  bar_alpha = 0.8,
  protein_darker = '#d94801',
  motif          = '#08519c'
)
# aliases
gc <- go_colors 
cl <- colors_ls
manual_fill   <- scale_fill_manual(values = colors)
manual_colors <- scale_color_manual(values = colors)

# Intention for future is to standardize, or take these as args and rename that way
# For plots we can take a lookup as input and then use the specific names for labs/guides
lookup <- c(
  obs_id = 'id_visit', 
  group = 'visit',
  symbol = 'protein',
  subject_id = 'subject.id',
  subject_id = 'sample_id'
)

# setwd is not working for me so prepend paths with ../ for now
files <- list(
  # match is old, prep py extraction bug
  # match = '../data/mining/240522-split/freq/match.csv',
  motifs = '../data/mining/240522-split/freq/matched_motifs.csv',
  var_levels = '../data/mining/240522-split/freq/var_levels.csv',
  expr_binned = '../data/mining/expression/expression_binned.csv',
  expr = '../data/zhong2021/expression.csv',
  meta = '../data/zhong2021/meta.csv',
  s4 = '../data/zhong2021/s4-plasma_protein_profiling.csv',
  s6 = '../data/zhong2021/s6-diff_proteins.csv')

params <- list()
params$thresh_abs_diff <- 40
```


ATM training/valid masks are hidden by taking the diff b/t all samples and the obs_ids found inside of the training graphs
```{r data table defs}
so()
train_obs <- function() 
  file.path(
    files$motifs %>% dirname %>% dirname,
    'train/graphs/') %>% 
  list.files(recursive = TRUE) %>%
  basename() %>%
  str_replace('.pkl', '') %>% 
  unique()

meta <- read_csv(files$meta) %>% 
  rename(any_of(lookup)) %>% 
  mutate(
    train_mask = case_when(
      obs_id %in% train_obs() ~ TRUE,
      .default = FALSE
    ),
    subject_id = as.character(subject_id)
  )

match <- read_csv(files$motifs) %>% 
  mutate_if(is.numeric, as.character) %>% 
  merge(select(meta, train_mask, obs_id))

sample.ids <- filter(meta, train_mask)$subject_id

# NB: create diffs based on training set only
diffs <- match %>% 
  filter(train_mask) %>% 
  diffs_table()

npx_sig <- read_csv(files$s6)$Protein
expression <- read_csv(files$expr_binned) %>% 
  rename_with(tolower) %>% 
  rename(any_of(lookup)) %>% 
  mutate(subject_id = as.character(subject_id))

npx_assembled <- expression %>% select(subject_id, group, symbol, npx)
```

# Describe Frequency
```{r}
dir.create('export/describe', showWarnings = FALSE)
```

```{r}
.pfx <- function(match){
    lapply(c('diff', 'abs_diff', 'freq'), function(x) {
      match %>% 
      diffs_table() %>% 
      mutate(freq = day0 + day14) %>% 
      ggplot(aes_string(x=x)) +
      geom_histogram(aes(y = after_stat(count)), binwidth=1, color='black', fill='white')
      # geom_density(aes(y = after_stat(count)), alpha=0.2, fill='#ff6666')
      }
      ) %>% plot_grid(plotlist=., ncol=1)
  }
pl <- list()
pl[['all']] <- .pfx(match)
i = 0
for (b in c(TRUE, FALSE)){
  g <- match %>% 
    filter(train_mask == b) %>% 
    .pfx()
  pl[[as.character(i)]] <- g
  i=i+1
}
g <- plot_grid(plotlist=pl, ncol=3, labels=c('All', 'Train', 'Test'))
g
save_plot('export/describe/motif-count-density-hist.png', g, base_asp = 1, base_height = 8)
```
The larges abs diff in the test set is about 10, 40 in training, 49 in pooled 

To be as true to the train/test split, I should only consider the training set when identifying which motifs to select.
- From train set: identify treshold >> get ids of the motif features
  - Therefore: *compute diffs on train set only*
- Filter the pooled set by these ids before using them for features analysis.
```{r}
.vec <- seq(35, 40, 1) 
.n <- lapply(.vec, function(x) filter(diffs, abs_diff >= x) %>% nrow)
data.frame(tresh_abs_diff = .vec, n_motif = unlist(.n)) %>% print
```


```{r}
lty <- 'dashed'

.pfx <- function(diffs, x) {
  # https://stackoverflow.com/questions/37802637/density-plots-with-counts
  diffs %>% 
    ggplot(aes_string(x=x)) +
    geom_histogram(aes(y = after_stat(count)), binwidth=1, color='black', fill='white')
    # geom_density(aes(y = after_stat(count)), alpha=0.25, fill='#ff6666')
}

p <- .pfx(diffs, 'diff') +
  geom_vline(xintercept = params$thresh_abs_diff, linetype=lty) +
  geom_vline(xintercept = -params$thresh_abs_diff, linetype=lty) +
  labs(
    y = 'Count',
    x = 'Group Difference',
    title = 'Motif Frequency'
    ) +
  scale_x_continuous(breaks = c(-params$thresh_abs_diff,0,params$thresh_abs_diff))
  
p
save_plot('export/describe/group-difference-eyeball-threshold.png', p)
```
# Dim Reduction Exploration
Need to replicate npx_assembled from before
npx assembled is tidy df with subject_id, group, var_id, npx(value float)
```{r}
```


Note: `bin_mots_for_tree(length(npx_sig))` here takes the filtered full dataset, and then
- recalculates the diffs(n_motifs = length(npx_sig)) according to the full dataset
  - n_motifs is slice_max(by=abs_diff, with_ties)
- filters the dataset by those motifs in the diffs ()
```{r}
so()

plot_cumvar <- function(){
  npx <- npx_assembled %>% 
    filter(symbol %in% npx_sig) %>% 
    npx_assembled_to_matrix()  

  mot <- match %>% 
    filt_thresh(diffs, parampbs$thresh_abs_diff) %>% 
    bin_mots_for_tree(length(npx_sig))

  dfs <- list(
    'protein' = npx,
    'motif'   = mot
  )
  res <- list()
  for (i in 1:length(dfs)) {
    name <- names(dfs)[[i]]
    res.pca <-  pca_fx(dfs[[i]])
    eig <- res.pca$eig
    
    res[[name]] <- 
      list(
        eig = eig,
        pcn = find_pcn(eig),
        U = res.pca$svd$U,
        name = name
      )
  }

  prep_plot <- function(eig, pcn) {
    eig %>% 
      data.frame() %>% 
      mutate(pcn = row_number()) %>% 
      dplyr::slice_head(n = 50)
  }

  x_len <- map_depth(res, 1, 'pcn') %>% unlist %>% max +1

  .pfx <- function(ls, fill){
    lcolor <-  'black'
    linetype <- 'dashed'
    linesize <- 0.8
    xlabs <- seq(0,50,5)
    ylabs <- c(0,25,50,75,95)
    
    df <- prep_plot(ls$eig, ls$pcn)
    
    df %>% 
      ggplot(aes(x=pcn, y=cumulative.percentage.of.variance, fill=fill)) + 
      geom_bar(stat = 'identity', just=0.5, 
              width = 1,
              color='black', fill=fill, alpha=0.8) +
      geom_hline(yintercept = 95,     linetype=linetype, linewidth=linesize) +
      geom_vline(xintercept = ls$pcn, linetype=linetype, linewidth=linesize) +
      labs(x='', y='') +
      scale_y_continuous(limits = c(0,100), labels = ylabs, breaks = ylabs,) + 
      scale_x_continuous( limits = c(0,x_len), labels = xlabs, breaks=xlabs,) +
      theme_minimal()
  }

  labsize = 10
  PALETTE <- c(cl['mot_blue'], cl['pro_oran']) # same palette as hc_sup
  # PALETTE <- get_palette2(2, 'Set1', 0) # same palette as rocs
  p <- plot_grid(
    NULL,
    .pfx(res$motif, PALETTE[1]),
    .pfx(res$protein, PALETTE[2]) +
      labs(x='', y='Cumulative Var %'),
    rel_heights = c(0.05,1,1), 
    labels = c('',
               paste('Motifs',   nrow(mot), 'x', ncol(mot)), 
               paste('Proteins', nrow(npx), 'x', ncol(npx))
              ),
    vjust=0.7, hjust=-0.05,
    label_size=12, label_fontface=2,
    ncol = 1 ) +
    draw_label('Principal Component Number', y=0.03, size=labsize, hjust=0.4 )

  list(
    plot = p,
    mot = mot,
    npx = npx,
    dim_red = res
  ) 
}

res <- plot_cumvar()
p <- res$plot
print(p)
save_plot('export/pca-eigen.pdf', p, base_asp = 1.2)
```

# Phenotype Classification
```{r}
so()
dir.create('export/rocs', showWarnings = FALSE)
```

### Debug
I am not getting the same npx results as in the project ../../3-extra-rocs. Investigate these files
```{r}
debug <- list()
debug[['file']] <- list(
  npx_assembled = "/home/nikoli/projects/00001-covid-motif-mining/data/2-motif-analysis/pipe/npx_assembled.csv",
  npx_file =  "/home/nikoli/projects/00001-covid-motif-mining/data/0-paper-data/s6-diff_proteins.csv",
  sample.ids = '/home/nikoli/projects/00001-covid-motif-mining/data/2-motif-analysis/split80/train/sample_ids.txt'
)

debug$data <- list(
  npx_assembled = read.csv(debug$file$npx_assembled) %>% 
    rename(any_of(lookup)),
  npx_sig = read.csv(debug$file$npx_file) %>% 
    rename(any_of(lookup)), 
  sample.ids = scan(debug$file$sample.ids, quiet=T)
)
debug$data$npx_all <- debug$data$npx_assembled %>%
  rename(any_of(lookup)) %>% 
  npx_assembled_to_matrix()

debug$data$npx_sig_mat <- debug$data$npx_assembled %>%
  rename(any_of(lookup)) %>%
  filter(symbol %in% get_sig_proteins(files$s6)) %>% 
  npx_assembled_to_matrix()
```

```{r}
.id_cols <- c('label', 'subject_id')
debug$data$npx_sig_mat %>%
  arrange(subject_id, label) %>% 
  select(sort(colnames(.))) %>%
  relocate(.id_cols) %>% 
  head()

npx_assembled %>% 
  filter(symbol %in% npx_sig) %>% 
  npx_assembled_to_matrix() %>% 
  arrange(subject_id, label) %>% 
  head()
```
#### subplot per pcn, but use debug npx data
```{r per pcn subplot comp, fig.height=5, fig.width=7}

.mca <- function(df, ncp){
  res <- df %>% 
    select(where(is.numeric)) %>% 
    mutate_if(is.numeric, as.factor) %>% 
    MCA(ncp = ncp, graph=FALSE)
  res$svd$U %>% 
    data.frame() %>% 
    cbind(select(df, !where(is.numeric)))
}

.pca <- reduce_dims

.get_rocs <- function(df, n_dims, dr_fx, sample.ids, prefix='', dr.only=FALSE){
  rocs <- list()
  if (!dr.only){
    # one w/o dim reduction
    roc <-  get_dr_feature(df, function(x){return(x)}, sample.ids)
    name <- paste(prefix, '\nAUC =', roc$auc)
    rocs[[name]] <- roc
  }
  for (pcn in n_dims) {
    roc <-  get_dr_feature(df, function(x){dr_fx(x, pcn)}, sample.ids)
    name <- paste(prefix, 'PCN =', pcn, '\nAUC =', roc$auc)
    rocs[[name]] <- roc
  }
  rocs
}

pfx <- function(rocs, palette, clab='', alpha=0.8){
  p <- rocs %>% 
    ggroc(
      aes = c('color'), 
      alpha    = alpha,
      size     = 1.5) +
    theme_classic(base_size = BASE_SIZE) +
    labs(color = clab)
  
  p
}

.run <- function(sample.ids){
  # These should be isolated into a function call for the whole cell
  # to avoid any careless namespace overriding
  feat_len <- length(npx_sig)
  
  npx_sig_mat <- debug$data$npx_assembled %>% 
    filter(symbol %in% get_sig_proteins(files$s6)) %>% 
    npx_assembled_to_matrix()
  
  mot <- match %>%
    bin_mots_for_tree(feat_len)
  
  n_dims <- c(15,30,45,50)

  # .prefix <- paste('len =', feat_len)
  .prefix <- ''
  
  .mr <- .get_rocs(mot, n_dims, .mca, sample.ids, .prefix,)
  .pr <- .get_rocs(npx_sig_mat, n_dims, .pca, sample.ids, .prefix)
  pl <- list()
  
  # plot_grid labels are hard to make nice, should use ether subplot titles or guide title to mark PCN
  subplot_labs <- c(
    paste('Input'), 
    paste('PCN:', n_dims)
  )
  
  for (i in 1:length(.mr)){
    .name <- function(roc){paste0('AUC = ', roc$auc)}
    subpl <- list()
    subpl[[1]] <- .pr[[i]]
    subpl[[2]] <- .mr[[i]]
    
    # Select which line goes on top
    bottom <- 2
    top <- 1
    color_labs <- c(.name(subpl[[1]]), .name(subpl[[2]]))
  
    pl[[i]] <- pfx(subpl, PALETTE) + 
      theme(
        axis.text = element_text(size=6),
        axis.title = element_text(size=8),
        legend.position=c(0.6, 0.4)) +
      labs(
        x='', y='',
        # title = subplot_labs[[i]]
      ) +
      guides(color=guide_legend(title=subplot_labs[[i]])) +
      scale_color_manual(
        values = PALETTE,
        labels = color_labs,
      )
  }
  
  pl[[length(.mr)]] <- pl[[length(.mr)]] + labs(x='Specificity', y='Sensitivity')
  
  pl
}  


BASE_SIZE <- 12
PALETTE <- c(cl[['mot_blue']], cl[['pro_oran']]) %>% unlist # same palette as hc_sup
PALETTE <- c(PALETTE[2], PALETTE[1])

pl <- .run(debug$data$sample.ids)

legend <- get_legend(
  pl[[1]] + 
    # DANGER These are hardcoded to assume that motifs are the second label (bottom auc)
    scale_color_manual(values=PALETTE, labels=c('Protein', 'Motif')) +
    guides(color=guide_legend(
        title='Feature',
        override.aes=list(alpha=1, size=4)
    )) +
    theme(
      # legend.key.size = unit(1, 'cm'),
      legend.text = element_text(size=12),
      )
) 

g <- plot_grid(plotlist=pl, ncol=3, labels = 'auto') +
  draw_grob(legend,
            0.27, -0.17,
            hjust=0, vjust=0)
g
```
### subplot by pcn
```{r per pcn subplot comp, fig.height=5, fig.width=7}
.mca <- function(df, ncp){
  res <- df %>% 
    select(where(is.numeric)) %>% 
    mutate_if(is.numeric, as.factor) %>% 
    MCA(ncp = ncp, graph=FALSE)
  res$svd$U %>% 
    data.frame() %>% 
    cbind(select(df, !where(is.numeric)))
}

.pca <- reduce_dims

.get_rocs <- function(df, n_dims, dr_fx, sample.ids, prefix='', dr.only=FALSE){
  rocs <- list()
  if (!dr.only){
    # one w/o dim reduction
    roc <-  get_dr_feature(df, function(x){return(x)}, sample.ids)
    name <- paste(prefix, '\nAUC =', roc$auc)
    rocs[[name]] <- roc
  }
  for (pcn in n_dims) {
    roc <-  get_dr_feature(df, function(x){dr_fx(x, pcn)}, sample.ids)
    name <- paste(prefix, 'PCN =', pcn, '\nAUC =', roc$auc)
    rocs[[name]] <- roc
  }
  rocs
}

pfx <- function(rocs, palette, clab='', alpha=0.8){
  p <- rocs %>% 
    ggroc(
      aes = c('color'), 
      alpha    = alpha,
      size     = 1.5) +
    theme_classic(base_size = BASE_SIZE) +
    labs(color = clab)
  
  p
}

.run <- function(sample.ids){
  # These should be isolated into a function call for the whole cell
  # to avoid any careless namespace overriding
  feat_len <- length(npx_sig)
  
  npx_sig_mat <- npx_assembled %>% 
    filter(symbol %in% get_sig_proteins(files$s6)) %>% 
    npx_assembled_to_matrix()
  
  mot <- match %>%
    bin_mots_for_tree(feat_len)
  
  n_dims <- c(15,30,45,50)

  # .prefix <- paste('len =', feat_len)
  .prefix <- ''
  
  .mr <- .get_rocs(mot, n_dims, .mca, sample.ids, .prefix,)
  .pr <- .get_rocs(npx_sig_mat, n_dims, .pca, sample.ids, .prefix)
  pl <- list()
  
  # plot_grid labels are hard to make nice, should use ether subplot titles or guide title to mark
  # PCN
  subplot_labs <- c(
    paste('Input'), 
    paste('PCN:', n_dims)
  )
  
  for (i in 1:length(.mr)){
    .name <- function(roc){paste0('AUC = ', roc$auc)}
    subpl <- list()
    subpl[[1]] <- .pr[[i]]
    subpl[[2]] <- .mr[[i]]
    
    # Select which line goes on top
    bottom <- 2
    top <- 1
    color_labs <- c(.name(subpl[[1]]), .name(subpl[[2]]))
  
    pl[[i]] <- pfx(subpl, PALETTE) + 
      theme(
        axis.text = element_text(size=6),
        axis.title = element_text(size=8),
        legend.position=c(0.6, 0.4)) +
      labs(
        x='', y='',
        # title = subplot_labs[[i]]
      ) +
      guides(color=guide_legend(title=subplot_labs[[i]])) +
      scale_color_manual(
        values = PALETTE,
        labels = color_labs,
      )
  }
  
  pl[[length(.mr)]] <- pl[[length(.mr)]] + labs(x='Specificity', y='Sensitivity')
  
  pl
}  

.post <- function(pl){
  legend <- get_legend(
    pl[[1]] + 
      # DANGER These are hardcoded to eassume that motifs are the second label (bottom auc)
      scale_color_manual(values=PALETTE, labels=c('Protein', 'Motif')) +
      guides(color=guide_legend(
          title='Feature',
          override.aes=list(alpha=1, size=4)
      )) +
      theme(
        # legend.key.size = unit(1, 'cm'),
        legend.text = element_text(size=12),
        )
  ) 
  
  g <- plot_grid(plotlist=pl, ncol=3, labels = 'auto') +
    draw_grob(legend,
              0.27, -0.17,
              hjust=0, vjust=0)
  g
}


BASE_SIZE <- 12
PALETTE <- c(cl[['mot_blue']], cl[['pro_oran']]) %>% unlist # same palette as hc_sup
PALETTE <- c(PALETTE[2], PALETTE[1])
```


```{r per pcn subplot comp, fig.height=5, fig.width=7} 
.run(sample.ids) %>% post()
```


### Examine effect of data split on protein ROCs
I've had a split where the input of npx_sig gives AUC = 1 which is wild.

```{r, eval=FALSE}
split_samples <- function(){
  t <- 1:50
  a <- sample(t, 40)
  a <- c(a, a) # Subject id appears twice (once per group)
  as.character(a)
}

for (i in 1:10) {
  .run(split_samples()) %>% 
    .post() %>% 
    print()
}
```


```{r per pcn subplot comp, fig.height=5, fig.width=7} 
# ROCS old split
.run(debug$data$sample.ids) %>% .post()
```
# Heirarchical Clustering


```{r}
so()
match %>% 
  filter(motif_id %in% motif_id_thresh(12, diffs)) %>% 
  motif_profile_bin() %>%
  select(1:5) %>%
  head
```


Too many motifs, not enough variation.
Try loosening the abs_diff threshold
Try dim red
Try calc diffs on the pooled data (IG, we are clustering the full pool anyways ..)
```{r fig.width=11, fig.height=11, eval=FALSE}
so()
mot <- match %>% 
  filt_thresh(., diffs_table(.), params$thresh_abs_diff)

sample.ids <- filter(meta, train_mask)$subject_id
  
runphylo(mot)
.saveplot('motif-all-dendro.png')

filter_ids(mot, sample.ids) %>%
  runphylo()
.saveplot('train-all-dendro.png')

filter_ids(mot, sample.ids, negative=TRUE) %>%
  runphylo()
.saveplot('test-all-dendro.png')
```

# MISC
