---
title: '' output: html_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
so <- function() {
  source('src.r')
  library(cowplot)
}
so()
```


```{r, include=FALSE}
## Theme ----
# don't know how to integrate this into a theme() call
colors    <- c('#BC0722', '#596f91')
go_colors <- c(colors, '#f16913', '#0DAC4B')
colors_ls <- list(
  disease        = colors[[1]],
  recovery       = colors[[2]],
  protein        = go_colors[[3]],
  prot_disease   = go_colors[[3]],
  prot_recovery  = go_colors[[4]],
  mot_blue = get_palette2(3, 'Blues', 4)[1],
  pro_oran = get_palette2(3, 'Oranges', 4)[1],
  bar_alpha = 0.8,
  protein_darker = '#d94801',
  motif          = '#08519c'
)
# aliases
gc <- go_colors 
cl <- colors_ls
manual_fill   <- scale_fill_manual(values = colors)
manual_colors <- scale_color_manual(values = colors)

# Intention for future is to standardize, or take these as args and rename that way
# For plots we can take a lookup as input and then use the specific names for labs/guides
lookup <- c(
  obs_id = 'id_visit', 
  group = 'visit',
  symbol = 'protein',
  subject_id = 'subject.id',
  subject_id = 'sample_id'
)

# setwd is not working for me so prepend paths with ../ for now
files <- list(
  match = '../data/mining/240522-split/freq/match.csv',
  motifs = '../data/mining/240522-split/freq/matched_motifs.csv',
  var_levels = '../data/mining/240522-split/freq/var_levels.csv',
  expr_binned = '../data/mining/expression/expression_binned.csv',
  expr = '../data/zhong2021/expression.csv',
  meta = '../data/zhong2021/meta.csv',
  s4 = '../data/zhong2021/s4-plasma_protein_profiling.csv',
  s6 = '../data/zhong2021/s6-diff_proteins.csv')
```

For initial results (ex. clustering, prediction performance) I don't need to regroup the var_levels in files$motifs back. But for GO analysis, this will be required.

There should be two files there: match (obs_id and motif id, makes the binary feature) and motifs (describes the motof patterns). Also maybe a third file, var_levels, expected. But today (240524) I am going to just proceed with classification first.

```{bash}
ls ../data/mining/240522-split/freq/
head ../data/mining/240522-split/freq/*
```


ATM training/valid masks are hidden by taking the diff b/t all samples and the obs_ids found inside of the training graphs
```{r}
train_obs <- 
  file.path(
    files$match %>% dirname %>% dirname,
    'train/graphs/') %>% 
  list.files(recursive = TRUE) %>%
  basename() %>%
  str_replace('.pkl', '') %>% 
  unique()

meta <- read_csv(files$meta) %>% 
  rename(any_of(lookup)) %>% 
  mutate(
    train_mask = case_when(
      obs_id %in% train_obs ~ TRUE,
      .default = FALSE
    ),
    subject_id = as.character(subject_id)
  )

# Here assembling the 'mm.file' from template.
match <- read_csv(files$match) %>% 
  rename(any_of(lookup)) %>% 
  mutate_if(is.numeric, as.character) %>% 
  merge(meta %>% select(subject_id, group, obs_id))
```

# Describe Frequency
```{r}
dir.create('export/describe', showWarnings = FALSE)
```

```{r}
.d <- diffs_table(match) %>% 
  mutate(freq = day0 + day14)

g <- lapply(c('diff', 'abs.diff', 'freq'), function(x) {.d %>% 
    ggplot(aes_string(x=x)) +
    geom_histogram(aes(y = after_stat(count)), binwidth=1, color='black', fill='white')
    # geom_density(aes(y = after_stat(count)), alpha=0.2, fill='#ff6666')
    }
    ) %>% 
  plot_grid(plotlist=., ncol=1)

g
save_plot('export/describe/motif-count-density-hist.png', g, base_asp = 1, base_height = 8)
```
The larges abs diff is 15. 

If I were eyeballing this one, for tails I'd set the diff threshold at 10. How many motifs does that give us? Quite a few.
Above 12 gives a compareable-ish number to the previous (very truncated) dataset

```{r}
.vec <- 11:13 
lapply(.vec, function(x) filter(.d, abs.diff > x) %>% nrow)

# lapply(.vec, function(x) {
#   filter(match, motif_id %in% motif_id_thresh(x, .d))$subject_id %>% unique %>% length
# })
```


```{r}
.pfx <- function(diffs, x) {
  # https://stackoverflow.com/questions/37802637/density-plots-with-counts
  diffs %>% 
    ggplot(aes_string(x=x)) +
    geom_histogram(aes(y = after_stat(count)), binwidth=1, color='black', fill='white')
    # geom_density(aes(y = after_stat(count)), alpha=0.25, fill='#ff6666')
}

thresh <- 12
lty <- 'dashed'
p <- .pfx(.d, 'diff') +
  geom_vline(xintercept = thresh, linetype=lty) +
  geom_vline(xintercept = -thresh, linetype=lty) +
  labs(
    y = 'Count',
    x = 'Group Difference',
    title = 'Motif Frequency'
    ) +
  scale_x_continuous(breaks = c(-thresh,0,thresh))
  
p
save_plot('export/describe/group-difference-eyeball-threshold.png', p)
```
# Dim Reduction Exploration
Need to replicate npx_assembled from before
npx assembled is tidy df with subject_id, group, var_id, npx(value float)
```{r}
npx_sig <- read_csv(files$s6)$Protein
expression <- read_csv(files$expr_binned) %>% 
  rename_with(tolower) %>% 
  rename(any_of(lookup)) %>% 
  mutate(subject_id = as.character(subject_id))

npx_assembled <- expression %>% select(subject_id, group, symbol, npx)
```


```{r}
so()

plot_cumvar <- function(){
  npx <- npx_assembled %>% 
    filter(symbol %in% npx_sig) %>% 
    npx_assembled_to_matrix()  

  mot <- match %>% 
    bin_mots_for_tree(length(npx_sig))

  dfs <- list(
    'protein' = npx,
    'motif'   = mot
  )
  res <- list()
  for (i in 1:length(dfs)) {
    name <- names(dfs)[[i]]
    res.pca <-  pca_fx(dfs[[i]])
    eig <- res.pca$eig
    
    res[[name]] <- 
      list(
        eig = eig,
        pcn = find_pcn(eig),
        U = res.pca$svd$U,
        name = name
      )
  }

  prep_plot <- function(eig, pcn) {
    eig %>% 
      data.frame() %>% 
      mutate(pcn = row_number()) %>% 
      dplyr::slice_head(n = 50)
  }

  x_len <- map_depth(res, 1, 'pcn') %>% unlist %>% max +1

  .pfx <- function(ls, fill){
    lcolor <-  'black'
    linetype <- 'dashed'
    linesize <- 0.8
    xlabs <- seq(0,50,5)
    ylabs <- c(0,25,50,75,95)
    
    df <- prep_plot(ls$eig, ls$pcn)
    
    df %>% 
      ggplot(aes(x=pcn, y=cumulative.percentage.of.variance, fill=fill)) + 
      geom_bar(stat = 'identity', just=0.5, 
              width = 1,
              color='black', fill=fill, alpha=0.8) +
      geom_hline(yintercept = 95,     linetype=linetype, linewidth=linesize) +
      geom_vline(xintercept = ls$pcn, linetype=linetype, linewidth=linesize) +
      labs(x='', y='') +
      scale_y_continuous(limits = c(0,100), labels = ylabs, breaks = ylabs,) + 
      scale_x_continuous( limits = c(0,x_len), labels = xlabs, breaks=xlabs,) +
      theme_minimal()
  }

  labsize = 10
  PALETTE <- c(cl['mot_blue'], cl['pro_oran']) # same palette as hc_sup
  # PALETTE <- get_palette2(2, 'Set1', 0) # same palette as rocs
  p <- plot_grid(
    NULL,
    .pfx(res$motif, PALETTE[1]),
    .pfx(res$protein, PALETTE[2]) +
      labs(x='', y='Cumulative Var %'),
    rel_heights = c(0.05,1,1), 
    labels = c('', 'Motifs', 'Proteins'),
    vjust=0.7, hjust=-0.05,
    label_size=12, label_fontface=2,
    ncol = 1 ) +
    draw_label('Principal Component Number', y=0.03, size=labsize, hjust=0.4 )

  p 
}

p <- plot_cumvar()
print(p)
save_plot('export/pca-eigen.pdf', p, base_asp = 1.2)
```

# Phenotype Classification
```{r}
so()
dir.create('export/rocs', showWarnings = FALSE)
```

### subplot by pcn
```{r per pcn subplot comp, fig.height=8, fig.width=6}
.mca <- function(df, ncp){
  res <- df %>% 
    select(where(is.numeric)) %>% 
    mutate_if(is.numeric, as.factor) %>% 
    MCA(ncp = ncp, graph=FALSE)
  res$svd$U %>% 
    data.frame() %>% 
    cbind(select(df, !where(is.numeric)))
}

.pca <- reduce_dims

.get_rocs <- function(df, n_dims, dr_fx, prefix='', dr.only=FALSE){
  rocs <- list()
  if (!dr.only){
    # one w/o dim reduction
    roc <-  get_dr_feature(df, function(x) return(x))
    name <- paste(prefix, '\nAUC =', roc$auc)
    rocs[[name]] <- roc
  }
  for (pcn in n_dims) {
    roc <-  get_dr_feature(df, function(x) dr_fx(x, pcn))
    name <- paste(prefix, 'PCN =', pcn, '\nAUC =', roc$auc)
    rocs[[name]] <- roc
  }
  rocs
}

pfx <- function(rocs, palette, clab='', alpha=0.8){
  p <- rocs %>% 
    ggroc(
      aes = c('color'), 
      alpha    = alpha,
      size     = 1.5) +
    theme_classic(base_size = BASE_SIZE) +
    labs(color = clab)
  
  p
}


feat_len <- length(npx_sig)
npx_sig_mat <- npx_assembled %>% 
  filter(symbol %in% npx_sig) %>% 
  npx_assembled_to_matrix()
mot <- match %>%
  bin_mots_for_tree(feat_len)

n_dims <- c(15,30,45,50)
BASE_SIZE <- 12
PALETTE <- c(cl[['mot_blue']], cl[['pro_oran']]) %>% unlist # same palette as hc_sup
PALETTE <- c(PALETTE[2], PALETTE[1])
# .prefix <- paste('len =', feat_len)
.prefix <- ''

.mr <- .get_rocs(mot, n_dims, .mca, .prefix)
.pr <- .get_rocs(npx_sig_mat, n_dims, .pca, .prefix)
pl <- list()

# plot_grid labels are hard to make nice, should use ether subplot titles or guide title to mark PCN
subplot_labs <- c(
  paste('Input'), 
  paste('PCN:', n_dims)
)

for (i in 1:length(.mr)){
  .name <- function(roc){paste0('AUC = ', roc$auc)}
  subpl <- list()
  subpl[[1]] <- .pr[[i]]
  subpl[[2]] <- .mr[[i]]
  
  # Select which line goes on top
  bottom <- 2
  top <- 1
  color_labs <- c(.name(subpl[[1]]), .name(subpl[[2]]))

  pl[[i]] <- pfx(subpl, PALETTE) + 
    theme(
      axis.text = element_text(size=6),
      axis.title = element_text(size=8),
      legend.position=c(0.6, 0.4)) +
    labs(x='', y='', 
         # title = subplot_labs[[i]]
    ) +
    guides(color=guide_legend(title=subplot_labs[[i]])) +
    scale_color_manual(
      values = PALETTE,
      labels = color_labs,
    )
}

pl[[length(.mr)]] <- pl[[length(.mr)]] + labs(x='Specificity', y='Sensitivity')

legend <- get_legend(
  pl[[1]] + 
    # DANGER These are hardcoded to assume that motifs are the second label (bottom auc)
    scale_color_manual(values=PALETTE, labels=c('Protein', 'Motif')) +
    guides(color=guide_legend(
        title='Feature',
        override.aes=list(alpha=1, size=4)
    )) +
    theme(
      # legend.key.size = unit(1, 'cm'),
      legend.text = element_text(size=12),
      )
) 
```

```{r per pcn subplot wide, fig.dim=c(7,5)}
g <- plot_grid(plotlist=pl, ncol=3, labels = 'auto') +
  draw_grob(legend,
            0.27, -0.17,
            hjust=0, vjust=0)

save_plot('export/rocs/roc-per-pcn-subplot-wide.pdf', g,
          base_asp = 7/5, base_height = 5)
g
```
# Heirarchical Clustering


```{r}
so()
match %>% 
  filter(motif_id %in% motif_id_thresh(12, .d)) %>% 
  motif_profile_bin() %>%
  select(1:5) %>%
  head
```


```{r fig.width=11, fig.height=11}
so()
mot <- match %>% 
  filter(motif_id %in% motif_id_thresh(12, .d))

mot$subject_id %>% unique %>% length

sample.ids <- filter(meta, train_mask)$subject_id

runphylo(mot)
.saveplot('motif-all-dendro.png')

filter_ids(mot, sample.ids) %>%
  runphylo()
.saveplot('train-all-dendro.png')

filter_ids(mot, sample.ids, negative=TRUE) %>%
  runphylo()
.saveplot('test-all-dendro.png')
```

# MISC
